\section{Appendix A - Glossary}
\label{sec:glossary}

The definitions provided below originate from~\cite{dp}.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{description}
%%%%%
\item[AE Partitioning with Dynamic VLAN] - Array Elements can be partitioned into
isolated groups by placing them on their own VLAN dynamically, meaning by remote
configuration.  This mechanism is used to enable the HILSE, and will also be used
to enable the WSU Test System during the WSU AIVC period and the WSU HILSE during
WSU Operations.
%%%%%
\item[Angular Resolution (AR)] - The point spread function  full-width at half-maximum associated with a given science product. In the context of ALMA Quality Assurance Level 2 (QA2), this refers specifically to the mean synthesized beam width of a specific science product from the representative spectral window (spw), generally a spectral cube or single-spw continuum image. It is one of the PI-specified performance parameters set at the Science Group (SG) level.
%%%%%
%\item[Archive and Pipeline Operations (APO) group] - APO is in charge of the data-processing infrastructure (processing nodes and disk space, the hardware infrastructure related to the ALMA Science Archive) and the associated databases.
%%%%%
\item[Array Element] - An antenna and its associated (per-antenna) Central Local Oscillator (CLO) modules located in the AOS Technical Building.
%%%%%
%\item[Astronomer on Duty (AoD)] - ALMA staff member from the JAO, ARC or ARC node who carries out observations on the telescope. 

%%%%%
%\item[ALMA Observing Tool (ALMA OT)] - Software used to set up observing projects for the ALMA Observatory. It can be used to generate and submit an observing proposal containing SGs (Phase 1) and to prepare and submit SGs and SBs for observation (Phase 2).
%%%%%
\item[ALMA Data Processing System (ALMA DPS)] - The new ALMA DPS, designed to meet the requirements of the WSU, will be formed from a combination of existing ALMA sub-systems and a new system being developed by the Data Management (DMS) Division of NRAO for ALMA (the Radio Astronomy Data Processing System Program; RADPS-ALMA) and the ngVLA. As the ALMA DPS, existing ALMA sub-systems and the new RADPS-ALMA will work together to provide the functionality required to process WSU data. Refer to [RD55] for more details. 

\item[ALMA Project Data Model (APDM)] - The information of any Observing Project is stored in the ALMA Archive following the ALMA Project Data Model (APDM). The APDM defines a series of generic structures that are shared by all Observing Projects, including the project status, the proposal information, the outcome of the proposal review process, and the Scheduling Blocks (SB)\index{Scheduling Block(s)}\seeonlyindex{SB}{Scheduling Block(s)} with their inter-dependencies.
All these structures are filled with the relevant information (or with links to it in the Archive) at the time of the creation of the project, and updated accordingly during its complete lifecycle.
%%%%%
\item[ALMA Quality Assurance Tool (AQUA)] - Web interface to assist DRMs, DRs and other staff members to search, review and modify the status of raw and processed data products during the quality assurance process.
%%%%%
\item[ALMA Regional Center (ARC)] - Centers at each of the three Executive regions (East Asia, Europe, and North America; EA, EU, and NA) staffed by scientists with expertise in radio astronomy, millimeter/submillimeter astronomy, and interferometry whose purpose is to provide the interface between the JAO and their respective regional scientific communities, provide operational support to the JAO, and to maximize the productivity of the ALMA telescope. 
%%%%%%
\item[ALMA Science Data Model (ASDM)] The final product from each observation, recorded in the Archive, is known as the ASDM\index{ALMA Science Data Model}. The ASDM\index{ALMA Science Data Model} is the ALMA implementation of the Science Data Model (SDM)\index{Science Data Model}\seeonlyindex{SDM}{Science Data Model}\Alsoindex{ALMA Science Data Model}{Science Data Model}\Alsoindex{Science Data Model}{ALMA Science Data Model}, which was developed nearly 20 years ago for ALMA and the Expanded VLA~\cite{Viallefond2006,Pokorny2009,Lucas2010}. 
Each ASDM\index{ALMA Science Data Model} has a unique name containing the Array and Execution identifiers as hexadecimal values. The ASDM\index{ALMA Science Data Model} contains the metadata (headers, descriptions of the observation setup, ancillary data, etc) and the binary data (the raw data itself). Details can be found in Chapters 5.8 and 12.2 of~\cite{c10_thb}.
%%%%%%
\item[Common Astronomy Software Applications - CASA] - The primary data processing software package used for ALMA and VLA data is CASA. One of its core functionalities is to support the calibration and imaging pipelines for ALMA, VLA, VLA Sky Survey, and the Nobeyama 45 m telescope.
%%%%%%
\item[Commissioning Scientist] - Scientists within ALMA who are involved in commissioning new observing modes, researching and writing requirements for heuristics, and validating the resulting features and the overall data processing system. %Many of these scientists are also members of the PLWG.
%%%%%
\item[Cube Analysis and Rendering Tool for Astronomy (CARTA)]\footnote{\url{https://cartavis.org/}} - Interactive program that allows the user to efficiently visualize, examine, and perform basic analysis on images and (hyper)cubes.
%%%%%
\item[Cycle] - ALMA's 1-yr observing period, usually starting in October. 
%%%%%
\item[Data Reducer (DR)] - Users internal to the ALMA project who are responsible for data processing in operations.
%%%%%%
\item[Data Reduction Manager (DRM)] - A staff role in the ALMA QA2 workflow. The DRM reviews the processing and science products of PI observations and sets the QA2 state. At the ARCs, DRMs typically advise data analysts or data reducers (DRs) and review their work, then make the final assessment and set the state. DRMs also perform other tasks related to ingesting the products (e.g. review of manually processed packages, and facilitating/troubleshooting the package upload and ingestion process). At the JAO, the role of DRM and data reducer is typically performed by a single person (a SACM). In the context of automated deliveries (``Fast-Lane''), the actions of the DRM are performed by a software process.
%%%%%
\item[Delivery] - Delivery is the process of making a given dataset available to the PI within the ALMA Science Archive and notifying the PI of its availability.
%%%%%
\item[DMS] - Data Management \& Software. A Division of NRAO, in charge of designing a data processing system for ALMA and ngVLA. More details can be found in Section 4 of~\cite{dp}. 
%%%%%
%\item[Data Management Group (DMG)] - DMG is in charge of the data processing, review, and delivery of the ALMA data, consists of experts in data reduction from the ARCs and the JAO.
%%%%%
\item[Dynamic Scheduling Algorithm (DSA)] - The scheduling subsystem is a tool that algorithmically determines the optimal SB to execute at any given moment based on a given set of criteria.
%%%%%
\item[Executive] - ALMA partner region (East Asia, Europe, and North America).
%%%%%
\item[Execution Block (EB)] - A single observational execution of an SB\index{Execution Block}\seeonlyindex{EB}{Execution Block}\Alsoindex{Scheduling Block(s)}{Execution Block}\Alsoindex{Execution Block}{Scheduling Block(s)}.
%%%%%
\item[Group Observing Unit Set (GOUS)] - The Group ObsUnitSets are the smallest group of observation from two or more distinct ALMA Arrays which need combination within a SG.
%%%%%
\item[Joint ALMA Observatory (JAO)] - The JAO provides the unified leadership and management of the operation of ALMA. The JAO is located in Santiago (Chile) and synchronizes activities of the Executives in East Asia, Europe, and North America, as well as at the ALMA site near San Pedro de Atacama, Chile. 
%%%%%
\item[Life Cycle] - The sequence of stages that a particular unit of data or a system goes through from its initial generation, capture, or creation to its eventual archival, deletion, or decommissioning at the end of its useful life.  ``Life Cycle'' (eXTended State System, XTSS) is an ALMA subsystem software component that handles the state of a project unit (project, SG, SB, EB etc.).  
%%%%%
\item[Manual Data Reduction (MDR) working group] - Science Operations group responsible for the data processing software need to process data manually, where ``manually'' refers to either not or not exclusively using the automated ALMA Pipeline.

%%%%%
\item[Measurement Set (MS)] - A database designed to hold radioastronomical data,\footnote{\url{https://casadocs.readthedocs.io/en/stable/notebooks/casa-fundamentals.html}} to be calibrated following the Measurement Equation approach by~\cite{Hamaker1996}.
% moved to RD list: \footnote{Hamaker, J.P.; Bregman, J.D.; Sault, R.J. 1996, A\&AS 117, 137}.
%%%%
%%%%%
\item[Member Observing Unit Set (MOUS)] - Member ObsUnitSets are the smallest structures within a GOUS, executed in a single array or configuration (e.g. TP, 7-m, or C-3).
\item[Metadata] - Data that provides information about other data, but not the content of the data itself. In current CASA table and ASDM\index{ALMA Science Data Model} parlance, metadata includes e.g. antenna, spw, observation subtables, but not information like visibilities in the actual data columns of the main table.
%%%%%
\item[Observing Unit Set (ObsUnitSet, OUS)] - An ObsUnitSet or OUS is subset of an ALMA observation that would result in a dataset that meets (part of) a PI-specified Science Goal and is the smallest unit that can be feasibly calibrated. For a more complete description of the ALMA project data structure, see Chapter 12 of~\cite{c10_thb}.
%%%%%
\item[Observing Tool (OT)] - The ALMA OT\index{Observing Tool} is a Java desktop application used for the preparation and submission of ALMA proposals (Phase 1) and, for those which are accepted, Scheduling Blocks\index{Scheduling Block(s)} preparation (Phase 2). 
%It is also used for preparing and submitting Director's Discretionary Time (DDT) proposals and Supplemental Call (ACA stand-alone) proposals. 
More information can be found at \url{https://almascience.eso.org/documents-and-tools/cycle10/alma-ot-quickstart}.
%%%%%
\item[Phase 2 Group (P2G)] - ALMA-wide group of staff who are responsible for preparing and modifying ALMA SBs. 
%%%%%
%\item[Pipeline Operator] - Individuals responsible for overseeing the hardware and software for Pipeline processing in operations.
%%%%%
\item[Pipeline Calibration Products] - A collection of (a) calibration tables, (b) applycal commands, (c) flagging tables, (d) auxiliary products (calibrator fluxes, antenna positions, etc.) and (e) scripts for restoring the calibration and to fully execute the pipeline again. 
%%%%%
\item[Pipeline Imaging Products] - Depending on the mitigations applied during the imaging process, the pipeline imaging products could include 
(a) a continuum image per spectral window and per science target,
(b) a continuum image, combining all spectral windows per science target,
(c) a continuum-subtracted line cube per spectral window and per science target, and
(d) a continuum-subtracted binned cube of the representative spectral window, for the representative science target.
%%%%%%%
\item[Pipeline Weblog (weblog)] - Collection of HTML documents generated by the ALMA Pipeline to show a summary of the characteristics of the dataset processed, along with diagnostic plots and QA scores related to the processing itself. Note that, for consistency, the term ``weblog'' is used throughout, but in future such reports may be changed to a different format than HTML. However, the concept of a report containing information and pngs detailing the data and processing information will remain the same.
%%%%%
\item[Pipeline Working Group (PLWG)] - Scientists and domain experts on the ALMA pipeline team who research and prototype heuristics, define and prioritize requirements for developments, perform validation at feature and end-to-end levels, write documentation, and support operations by answering questions and investigating reported problems. Refer to~\cite{dp,dpda} for more details. 
%%%%%
\item[Project Tracker (ProTrack)] - The software tool that allows ALMA staff to track the observational and processing status of ALMA science projects.
%%%%%%
\item[Principal Investigator (PI)] - Individuals who submit observing proposals for ALMA data.
%%%%%
\item[Project] - A PI observing project that results from an accepted observing proposal. Projects consist of a set of SGs.
%%%%%
\item[Quality Assurance Zero (QA0)] - QA0 is the first quality assurance step after an SB is observed (referred to as an ``execution'') with the goal of determining whether the resulting data sets are valid, calibratable and contribute to the completion of the MOUS science objective. This is done soon after an observation is completed. The possible QA0 statuses are Pass, Semi-pass and Fail. QA0=Semi-pass is used for datasets that do not fulfill all the QA0 criteria, but contain data that is deemed of some scientific value (for instance, the presence of usable bandpass/amplitude calibrator data). Refer to Chapter 11.2 of~\cite{c10_thb} for details.
%%%%%%
\item[Quality Assurance One (QA1)] - QA1 is the quality assurance process that tracks the array and antenna performance parameters which vary slowly (typically on scales longer than a week) and which can affect data quality. Refer to Chapter 11.4 of~\cite{c10_thb} for details. 
%%%%%%%%
\item[Quality Assurance Two (QA2)] - QA2 is the quality assurance process that verifies whether the data that have been acquired meet the PI-specified performance parameters. The possible QA2 statuses are Pass, resulting in ingestion of products to the Archive and delivery to the PI; Semi-pass, meaning applicable performance or quality parameters were not met but will be ingested and delivered anyway; or Fail, which sends the SB back to the telescope for more observations to be collected. Refer to Chapter 11.5 of~\cite{c10_thb} for details.
%%%%%%%%
\item[Quality Assurance Three (QA3)] -  QA3 is the process that is engaged to diagnose and fix data quality issues identified after ingestion to the Archive and delivery to the PI.  Refer to Chapter 11.6 of~\cite{c10_thb} for details. 
%%%%%%
\item[Resubmission] - A previously-approved PI observing project that is submitted again to a new proposal in a subsequent observing cycle in order to complete the originally proposed observations.
%%%%%%%
\item[Scan] - In accordance with the description in the ALMA Software Science Requirements~\cite{SSRdoc}, a Scan is a time interval of data acquisition during which a single target source is observed with a particular Spectral Specification and visibility integration time. Scans are assembled from one or more subscans, where a subscan contains at least one visibility integration time.
%%%%%%%%%
\item[Scheduling Block (SB)] - A Scheduling Block\index{Scheduling Block(s)} is the smallest observing unit in ALMA that can be scheduled independently and contains a full description of how the science and the calibration targets have to be observed. Each run of an SB produces an Execution Block (EB). Refer to Chapter 12.1 of~\cite{c10_thb} for details.
%%%%%
\item[Science Goal (SG)] -  Science Goals are set up in the ALMA Observing Tool and are designed to capture the scientific aims of a particular observation whilst hiding the technical details of the observations from the user. The technical information is contained within SBs that are automatically generated from the Phase-2 SG. A SG consists of one or more SBs.
%%%%%
\item[Science Goal Observing Unit Set (SOUS)] - The top-level ObsUnitSets are called `Science Goal ObsUnitSets’; they correspond to the individual SGs of an Observing Project. 
%%%%%
\item[Session] - Group of EBs observed as the continuous execution of a single SB, until a certain scientific goal is reached, while keeping the SB information in run-time memory for the sake of observation integrity.
%%%%%%
\item[Software System Requirements (SSR)] - Control software that generates an on-the-fly sequence of commands and structure necessary for the actual execution of an SB on the telescope at observing time. Part of the ONLINE software subsystem.
%%%%%%
\item[Spectral Window (spw)] - A continuous spectral region with uniform spectral resolution to be observed, as set by the PI in the ALMA Observing Tool (ALMA OT).
%%%%%
\item[Stage] - Logical processing step e.g. ``solve bandpass'', or ``create one type of image''. Details will need to be refined in future, but for the purposes of this document refers to items with scope similar to current Pipeline tasks.
%%%%%
\item[Subscan] - A subdivision of a Scan.
All subscans within a Scan must be encoded with the same scan intent, corresponding to either the science target or one of the calibration intents (such as atmosphere, phase, bandpass, pointing, flux, polarization). Subscans may differ in their subscan intent, for example, atmospheric (\Tsys) calibrations are performed in a single scan with individual subscans acquired at different positions of the ALMA calibration device: hot load, ambient load, on-source, and sky. Subscans within a Scan may differ in their phase center, for example, pointed mosaics are observed using subscans in which different pointing positions within the source (called fields) correspond to independent subscans within a Scan.
%%%%%%%%
\item [Telescope Monitoring \& Control Database (TMCDB)] - It is used to store information on the state of various devices and restore the system after restarts. It also allows one to dynamically change values of the parameters to be used more or less on the fly. Some of the most commonly consulted and updated portions of TMCDB are the telescope pointing and focus models, pad locations, tilts and cable delays, and the delay models.
%%%%%%%%
\item[Water Vapor Radiometer (WVR)] - Device designed to track path length variations due to atmospheric water vapor. It generates data used to compute phase corrections to increase the coherence of interferometric ALMA data.
%%%%%
\item[WSU Test System] - The use of AE Partitioning with Dynamic VLAN to establish a group
of antennas for AIVC testing on ATAC\index{Advanced Technology ALMA Correlator} or TPGS\index{Total Power GPU Spectrometer} via the Parallel operations mode of Parallel Deployment.  These antennas will have been retrofitted with the WSU signal chain, and the ALMA Production Environment associated with them will be running a different version of software from the legacy ALMA system. 
\end{description}

\begin{comment}
\textbf{Aos Check}:

\textbf{Bulk Data Handling}:

\textbf{Offline System}:

\textbf{TelCal\index{TelCal}}:

\textbf{ALMA Science Archive Query Interface}

\textbf{Data Packer}

\textbf{Data Tracker}

\textbf{Harvester}

NGAS Archive 

Oracle Archive

Product Ingestor

Request Handler

TMCDB

Data Capturer

Quick Look

Cluster/ADAPT Interface (CLAI)

Project Tracker
\end{comment}

